Approximation as Differentiation: Notes

### 2025-05-25 main/category-of-families.tex
Prob ok for first pass on forward-mode AD.

We can verify that $\tangents$ is functorial, regardless of whether the linear maps $\RR^m \linearto \RR^n$
happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing composite forward
derivatives automatically so long as the primitive operations of the system are appropriately equipped with
them.

We can see these maps as living in $\Fam(\FinVect)$ if we regard every finite vector space $\RR^n$ as paired
with the family of tangent spaces which is constantly $\RR^n$ (also written $\RR^n$, for convenience), and
each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$ of a differentiable
function $f_1: \RR^m \to \RR^n$ and family of linear maps $f_2: \RR^m \to \RR^m \linearto \RR^n$. Then the
composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp \partial f):
(\RR^m, \RR^m) \to (\RR^k, \RR^k)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial
g}{f} \comp \partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ computing
the pushforward.

>>>>

We can verify that $\tangents$ is functorial, regardless of whether the linear maps $\RR^m \linearto \RR^n$
happen to be derivatives. But if $f_2(x)$ is the derivative of $f_1$ at $x$ for any $x \in \RR^m$ and
similarly for $g$ then the derivatives compose according to the forward chain rule and $(g \comp f)_2(x)$ is
the derivative of $(g \comp f)_1$ at $x$. Thus this setup provides a way of computing the forward derivative
of a composite function, without have to recompute the function itself, as long as all atomic operations
provide derivatives. An efficient implementation of \GPS based on $\Fam(\LatGal)$ (\secref{fam:galois-slicing}
above) would most likely apply a similar optimisation to make it possible to compute forwards slices without
having to reexecute the program at the same time.

We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$, with the reindexing along $f$ selecting the appropriate
pushforward by recomputing $f(x)$.
